---
title: "Analyses-Detect-Pulses"
author: "Andrew Barr"
date: "September 28, 2015"
output: 
  html_document:
    keep_md: true
---

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8, cache=TRUE)
library(parallel)
library(ggplot2)
library(phytools)
library(paleotree)
library(NullTurnover)
theme_set(theme_bw(20))
```


## Does sample size change the number of pulses detected? 

```{r}
sampSizes <- ceiling(seq(from=15, to=300, by = 15))
sampSizeResults <- lapply(sampSizes, FUN=function(x){
  
  pulses <- mclapply(1:1000, mc.cores=4, FUN=function(y) {
    detectPulses(nTaxa = x, desiredBinNumber = 15, plotRates = FALSE)
  })
  return(mean(unlist(pulses)))
})
```

```{r}
qplot(sampSizes, unlist(sampSizeResults), size=I(4)) +   
  labs(y="mean number of pulses", title="mean number of pulses by sample size", x="sample size")
```

## Yes....larger sample sizes converge to a value...what about variablilty?

```{r}
sampSizes <- ceiling(runif(1000, min=10, max=200))
sampSizeResults <- lapply(sampSizes, FUN=function(x){

    detectPulses(nTaxa = x, desiredBinNumber = 15, plotRates = FALSE)
 
})

```

```{r}
qplot(factor(unlist(sampSizeResults)), sampSizes, size=I(4), geom="violin") +  
  labs(x="number of pulses", title="mean number of pulses by sample size", x="sample size")
```

## Does bin # change the number of pulses detected? Yes....positively and linearly

```{r}
binNumbers <- seq(from=4, to=30, by=2)
binNumberResults <- lapply(binNumbers, FUN=function(x){
  
  pulses <- mclapply(1:1000, mc.cores=4, FUN=function(y) {
    detectPulses(nTaxa = 50, desiredBinNumber = x, plotRates = FALSE)
  })
  return(mean(unlist(pulses)))
})

```

```{r}
summary(lm(unlist(binNumberResults) ~ binNumbers))
```

```{r}
qplot(binNumbers, unlist(binNumberResults), size=I(4)) +   
  labs(y="mean number of pulses", title="mean number of pulses by number of bins", x="number of bins") + 
  stat_smooth(method="lm", se=F, shape=2) + 
  annotate(27, 2, geom="text", label="r^2==0.99", parse=T)
```

## Does extinction rate matter?


```{r}
exRates <- seq(from=0, to=1, by = 0.05)
exRateResults <- mclapply(exRates, mc.cores=4, FUN=function(x){
  
  pulses <- lapply(1:1000, FUN=function(y) {
    detectPulses(nTaxa = 75, desiredBinNumber = 15, deathRate = x, plotRates = FALSE)
  })
  return(mean(unlist(pulses)))
})

```

```{r}
qplot(exRates, unlist(exRateResults), size=I(4)) +    
  labs(y="mean number of pulses", title="mean number of pulses by extinction rate", x="extinction rate" )
```

```{r}
extinctionRates <- runif(200)

extinctionRateResults <- lapply(extinctionRates, FUN=function(x){
  detectPulses(nTaxa = 50, desiredBinNumber = 15, plotRates = FALSE)
})

```

```{r}
qplot(factor(unlist(extinctionRateResults)), extinctionRates, size=I(4), geom="violin") +   
  labs(x="number of pulses", title="number of pulses by extinction rate", y="number of pulses")
```

## Sanity check, to make sure that the criterion works like I want it to

```{r}
criteria <- seq(from=1, to=3, by=0.2)
criteriaResults <- lapply(criteria, FUN=function(x){
  
  pulses <- mclapply(1:1000, mc.cores=4, FUN=function(y) {
    detectPulses(nTaxa = 50, desiredBinNumber = 10, criterion = x, plotRates = FALSE)
  })
  return(mean(unlist(pulses)))
})
```

```{r}
qplot(criteria, unlist(criteriaResults), size=I(4)) +    
  labs(y="mean number of pulses", title="mean number of pulses by pulse criterion", x="pulse criterion")
```

```{r}
criteria <- runif(200, min=1, max=2.5)

criteriaResults <- lapply(criteria, FUN=function(x){
  detectPulses(nTaxa = 50, desiredBinNumber = 15, criterion = x,  plotRates = FALSE)
})

```

```{r}
qplot(factor(unlist(criteriaResults)), criteria, size=I(2), geom="violin") +   
  labs(x="number of pulses", y="pulse criterion", title="number of pulses by criterion")
```

## How often do you get pulses with parameters like Plio-Pleistocene african bovids?

```{r}
library(readxl)
bibi <- read_excel("/Users/wabarr/Dropbox/TurnoverPulseRedux/Bibi-Kiessling-PNAS-Supp1.xlsx")

inRange <- function(lowerBound, upperBound, FAD, LAD) {
  ifelse(LAD > lowerBound && FAD < upperBound, return(TRUE), return(FALSE))
}

#how many taxa are in the range 3.75 - 1.25?

taxaInRange <- sapply(1:nrow(bibi), FUN=function(rowIndex) {inRange(lower=1.25, upper=3.75, FAD=bibi[rowIndex,"FAD"], LAD=bibi[rowIndex,"LAD"])})
sum(taxaInRange)

# what is the extinction rate in that interval, with respect to the originiation rate?
sum(((bibi$LAD < 3.75) * (bibi$LAD > 1.25))) / sum(((bibi$FAD < 3.75) * (bibi$FAD > 1.25)))


bibiResults <- sapply(1:1000, FUN=function(x) detectPulses(treeDepth = 2.5, criterion = 1.5, nTaxa=52, desiredBinNumber = 10, deathRate =  0.8533333))

```

```{r}
sum(bibiResults > 0) / length(bibiResults)
qplot(factor(bibiResults), geom="bar", position="stack") +  
  labs(x="number of pulses") + 
  annotate(x=2.5, y=600, geom="text", label=paste0(100*round(sum(bibiResults > 0) / length(bibiResults),2),"% of simulations \ndetect at least one pulse"), size=8) + 
  labs(title="Number of pulses: African bovid parameters")

```

